---
title: "PeerAssessment"
author: "Vivek Yadav"
date: "Tuesday, September 16, 2014"
output: html_document
---

# Overall steps
Below are steps to tidy data from the raw files. I have included R-code also for documentation. The same code is available for download in run_analysis.R file. 

1. Read data sets
2. Read features 
3. Read activity labels, and assign activity labels to corresponding activity code
4. Get features containing mean and std of variables. Included mean of frequency do main variables also.
5. Create test and training data sets.
6. Merge data sets.
7. Take aggregate
8. Remove redundant columns and write to table

```{r message=FALSE, eval=FALSE}
library(dplyr)
```
# 1: Read data sets
## reading test data

```{r message=FALSE, eval=FALSE}
xtest = read.table('./UCI HAR Dataset/test/X_test.txt', 
                   header = FALSE,stringsAsFactors = FALSE)
ytest = read.table('./UCI HAR Dataset/test/Y_test.txt', 
                   header = FALSE,stringsAsFactors = FALSE)
subtest = read.table('./UCI HAR Dataset/test/subject_test.txt', 
                      header = FALSE,stringsAsFactors = FALSE)
```

## reading train data  
```{r message=FALSE, eval=FALSE}
xtrain = read.table('./UCI HAR Dataset/train/X_train.txt', 
                   header = FALSE,stringsAsFactors = FALSE)
ytrain = read.table('./UCI HAR Dataset/train/Y_train.txt', 
                   header = FALSE,stringsAsFactors = FALSE)
subtrain = read.table('./UCI HAR Dataset/train/subject_train.txt', 
                      header = FALSE,stringsAsFactors = FALSE)
```
# 2:reading features

```{r message=FALSE, eval=FALSE}
features = read.table('./UCI HAR Dataset/features.txt', 
                    header = FALSE,stringsAsFactors = FALSE)
```

# 3: reading Activity labels
```{r message=FALSE, eval=FALSE}
ActivityLabels = read.table('./UCI HAR Dataset/Activity_labels.txt', 
                      header = FALSE,stringsAsFactors = FALSE)
ActivityLabels_train = subtrain
for (i in ActivityLabels$V1) { ActivityLabels_train[i==ytrain] =  ActivityLabels$V2[i] }

ActivityLabels_test = subtest
for (i in ActivityLabels$V1) { ActivityLabels_test[i==ytest] =  ActivityLabels$V2[i] }
```

# 4: getting features with mean and std. Took mean and std before to store fewer numbers.
```{r message=FALSE, eval=FALSE}
mean_ind<-grep("mean()",features$V2) # did not use fixed = TRUE to include mean of frequency data also.
std_ind<-grep("std()",features$V2)
```
# 5a: creating test data
```{r message=FALSE, eval=FALSE}
## added 4 new columns to quantify subject, Activity, ActivityLabel, and dataType (test/train)
data_test<- data.frame(subtest,ytest,ActivityLabels_test,
                       rep("test", times = dim(ytest)[1]),xtest[,mean_ind],
                       xtest[,std_ind])
```
## adding column names
```{r message=FALSE, eval=FALSE}
colnames(data_test)<-c("subject","Activity","ActivityLabel"
                       ,"dataType",features$V2[mean_ind],
             features$V2[std_ind])
```

# 5b: creating train data
```{r message=FALSE, eval=FALSE}
data_train<- data.frame(subtrain,ytrain,ActivityLabels_train,
                        rep("train", times = dim(ytrain)[1]),
                        xtrain[,mean_ind],
                       xtrain[,std_ind])
```

## adding column names
```{r message=FALSE, eval=FALSE}
colnames(data_train)<-c("subject","Activity","ActivityLabel"
                        ,"dataType",features$V2[mean_ind],
                       features$V2[std_ind])
```
# 6: Merging data sets data
```{r message=FALSE, eval=FALSE}
data_all<-rbind(data_test,data_train) # dataset with test and train data combined.
```
# 7: taking aggregate
```{r message=FALSE, eval=FALSE}
aggdata <-aggregate(data_all, by=list(subject = data_all$subject , Activity =  data_all$Activity,
                                      ActivityLabel =  data_all$ActivityLabel,
                                      dataType = data_all$dataType), 
                                      FUN=mean,rm.na = T)
```

# 8: removing redundant columns and write
```{r message=FALSE, eval=FALSE}
aggdata<-aggdata[,-c(5:8)]
write.table(aggdata,"TidyAveragedActivityData.txt",row.name=FALSE)
```

# read using 
```{r message=FALSE, eval=FALSE}
data_agg<- read.table('TidyAveragedActivityData.txt',header=TRUE)
```